The file CA2.csv contains 100 observations on 12 unknown variables. Consider this as some data matrix X. Using Singular Value Decomposition, find lower rank approximations of X for all ranks from 1 – 12

```{r}
X <- read.csv("CA2.csv")
```

```{r}
s <- svd(X)
svd_approx <- function(rank){
  app <- s$u[,1:rank] %*% as.matrix((diag(s$d))[1:rank, 1:rank]) %*% t(s$v[, 1:rank])
  return(app)
}
X_k <- list()
for (i in 1:12){
  X_k[[i]] <- svd_approx(i)
}
X_k[[1]]
```

# Question 1

For each approximation $\tilde{X}_k$ of rank k, calculate the error, $\Delta _k=X-\tilde{X}_k$

```{r}
delta <- X - X_k[[4]]
mean_delta <- colMeans(delta)
as.matrix(mean_delta)
```

------------------------------------------------------------------------

# Question 2

Compare the correlation matrix of X with that of $\tilde{X}_2$and briefly interpret.

```{r}
library(ggcorrplot)
x_cor <- round(cor(X), 1)
ggcorrplot(x_cor, "square", lab = TRUE)

x_2_cor <- round(cor(X_k[[2]]), 1)
ggcorrplot(x_2_cor, "square", lab =TRUE)
```

------------------------------------------------------------------------

# Question 3

Calculate the Frobenius norm, defined as

$$
 ||A||_F= \sqrt{\sum_{i=1}^{m}\sum_{j=1}^{m}|aij|^2}
$$

for $\Delta_k$, k =1,2.....12. Plot the Frobenius norm as a function of k and briefly describe your findings.

```{r}
X_k <- list()
for(i in 1:12){
  X_k[[i]] <- svd_approx(i)
}

FN <- c()
Frobenius_Norm <- function(rank){
   FN[rank]<- sqrt(sum(abs(X_k[[rank]]^2)))
}

for(i in 1:12){
  FN[i] <- Frobenius_Norm(i)
}

df <- data.frame(x = 1:12, y = FN)
library(ggplot2)
ggplot(df, aes(x=x, y=y))+
  geom_line()
```

\

------------------------------------------------------------------------

# Question 4

```{r}

```
