The file CA2.csv contains 100 observations on 12 unknown variables. Consider this as some data matrix X. Using Singular Value Decomposition, find lower rank approximations of X for all ranks from 1 – 12

```{r}
#| echo: false
dat <- read.csv("CA2.csv")
```

```{r}
#| echo: false
library(Matrix)
s <- svd(dat)
scd_approx <- function(rank){
  app <- s$u[,1:rank] %*% as.matrix(diag(s$d)[1:rank, 1:rank]) %*% t(s$v[,1:rank])
  app
}
```

# Question 1

```{r}
#| eval: false
library(knitr)
X <- as.matrix(dat)
delta_4 <- X - scd_approx(4)
mean_error_vector <- colMeans(delta_4) |> round(4)
kable(mean_error_vector)
```

# Question 2

```{r}
<<<<<<< HEAD
library(ggcorrplot)
library(patchwork)

x_cor <- round(cor(X), 1)
p1 <- ggcorrplot(x_cor, "square", lab = TRUE, title = "Original X Matrix")

x_2_cor <- round(cor(X_k[[2]]), 1)
p2 <- ggcorrplot(x_2_cor, "square", lab =TRUE, title = "Approximated X Matrix Using \nRank 2")

p1 + p2
```

**Interpretation:**

1.  The original data shows that most of the variables are uncorrelated, the rank 2 approximation shows that they are corerlated. For example with V6 and V8 in the original matrix are uncorrelated while the rank 2 approximation correlation coefficient is 0.6.
2.  The original data shows slight correlations between variables with each other while the rank 2 approximation of X exagerates these correaltions. For example with V12 and V11 in the original matrix is -0.2 while the rank 2 approximation is -0.9.
3.  The rank 2 captures the 2 largest singular values and therefore attempts to capture a majority of the variation in the original X matrix but discards the other higher order variations. Therefore the rank 2 approximation has altered some of the structures in the data and has affected the relationships between variables, changing their correlation coefficients.

------------------------------------------------------------------------

# Question 3

Calculate the Frobenius norm, defined as

$$
 ||A||_F= \sqrt{\sum_{i=1}^{m}\sum_{j=1}^{m}|aij|^2}
$$

for $\Delta_k$, k =1,2.....12. Plot the Frobenius norm as a function of k and briefly describe your findings.

```{r}
FN <- c()
Frobenius_Norm <- function(rank){
   FN[rank]<- sqrt(sum(delta[[rank]]^2))
}

for(i in 1:12){
  FN[i] <- Frobenius_Norm(i)
}

df <- data.frame(x = 1:12, y = FN)
library(ggplot2)
ggplot(df, aes(x=x, y=y))+
  geom_line() +
  labs (
    title = "Frobenius Norm Plot For Delta Error of \n
    Rank k = 1...12 approximations of X", 
    x = "Rank k", 
    y = "Frobenius Norm"
  )
```

**Interpretation:**

1.  Decreasing (almost linear) trend of the Frobenius Norm as the the approximation of rank K increases.
2.  In other words, the delta error between the original matrix and the approximated rank K matrix decreases as the approximation gets closer to rank 12, where the original data set and the approximation are equal and the error is zero. \

------------------------------------------------------------------------

# Question 4

Plot the percentage of the total variation in X retained in $\tilde{X}_k$ for K =1,...12. Again, briefly interpret.

```{r}
total_var <- sum(s$d^2)
retained_var <- c()

for (i in 1:12){
  retained_var[i] <- ((sum(s$d[1:i]^2)) / total_var) * 100 
}

df <- data.frame(x = 1:12, y = retained_var)
library(ggplot2)
ggplot(df, aes(x=x, y=y))+
  geom_line() +
  labs (
    title = "Percentage of the total variation in X
    that is retained for each rank K approximation of X", 
    x = "Rank k", 
    y = "Percentage of Variance Retained"
  )
=======
#| eval: false
X2_approx <- scd_approx(2)
corr_X <- cor(X)
corr_X2 <- cor(X2_approx)
kable(round(corr_X, 4), caption = "Correlation Matrix of X")
kable(round(corr_X2, 4), caption = "Correlation Matrix of Rank-2 Approximation")
corr_diff <- abs(corr_X - corr_X2)
kable(round(corr_diff, 4), caption = "Absolute Differences in Correlation Matrices")
>>>>>>> 1026bac374c4905fc898767296ffb6bc938a4668

print(paste("Percentage of variance retained rank 7:", round(retained_var[8],2)))
```

-   Some values in the table are quite large (e.g., 0.9768 for V4, V9 and 0.9541 for V7, V9), showing that the correlation between those variable pairs has significantly changed in the low-rank approximation. This suggests that the rank-2 approximation fails to retain relationships in the data.

<<<<<<< HEAD
1.  The amount of variation in the original matrix retained increasing at a decreasing rate by the rank k approximation as the the approximation of rank K increases.

2.  This makes sense as we use more eigenvalues to approximate our original matrix, we keep more and more of the variation in the original matrix.

3.  Shows that by the time the rank k = rank 8, the approximation basically retains 92% of the variation of in the original matrix. Therefore using rank 8 approximation captures most of the data structures in the original matrix while being in lower dimensions and removing some of the linear dependencies between variables.
=======
-   Some values are close to 0 (e.g., 0.0197 for V6, V9), meaning the rank-2 approximation preserves these relationships well.

<!-- -->

-   The closer these values are to zero, the more accurately the rank-2 approximation maintains the correlation structure.
>>>>>>> 1026bac374c4905fc898767296ffb6bc938a4668
